---
title: "Pollinator graphs"
author: "Weinberg"
dat01e: "2023-11-30"
output:
  pdf_document: default
  html_document: default
editor_options: 
  markdown: 
    wrap: 72
---

#Box 0. Run this whole chunk to load the libraries

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!require("pacman")) install.packages("pacman")
  pacman::p_load(tidyverse, ggplot2, readxl, RColorBrewer)
```

# Box 1. Import the data


```{r}
#Change the name in quotation marks so that it matches the filename of your datasheet EXACTLY
setwd("~/Alice-Analysis/Alice_Analysis")
dat01<-read_excel("FINAL LAB_Marigolds Real+Artificial+Insect Behavior Data.xlsx")

```
#Box 2.
Tables of summary statistics. The tables show (in order)
1. Mean and standard error of the number of observations of each behavior by time
2. Mean and standard error of the number of observations of each behavior by flower type
3. Mean and standard error of the number of observations of each insect by time of day
4. Mean and standard error of the number of observations of each behavior by time of day


```{r message=FALSE}
se=function(x){sd(x)/sqrt(length(x))}

#Behaviors by time of day
SumStats_behavior_time<-dat01 %>% pivot_longer(colnames(dat01[6:ncol(dat01)]), names_to="Insect", values_to="Observations") %>% group_by(Time, Behavior) %>% summarise(Average=mean(Observations), SE=se(Observations))

SumStats_behavior_time

#behaviors by flower type (real or artificial)
SumStats_behavior_flower<-dat01 %>% pivot_longer(colnames(dat01[6:ncol(dat01)]), names_to="Insect", values_to="Observations") %>% group_by(Flower, Behavior) %>% summarise(Average=mean(Observations), SE=se(Observations))

SumStats_behavior_flower

#type of insect by time of day
SumStats_taxa_time<-dat01 %>% pivot_longer(colnames(dat01[6:ncol(dat01)]), names_to="Insect", values_to="Observations") %>% group_by(Time, Insect) %>% summarise(Average=mean(Observations), SE=se(Observations))

SumStats_taxa_time

SumStats_taxa_flower<-dat01 %>% pivot_longer(colnames(dat01[6:ncol(dat01)]), names_to="Insect", values_to="Observations") %>% group_by(Flower, Insect) %>% summarise(Average=mean(Observations), SE=se(Observations))
SumStats_taxa_flower
```

#Box 3. Plots of insects during daytime and nighttime
The plots are as follows:

1.Behaviors observed by flower (daytime)
2.Behaviors observed by flower (evening)
3. Insects observed by flower (daytime)
4. Insects observed by flower (evening)


```{r message=FALSE}
#summary statistics grouped by time of day and flower type
SumStats_behavior<-dat01 %>% pivot_longer(colnames(dat01[6:ncol(dat01)]), names_to="Insect", values_to="Observations") %>% group_by(Flower, Time, Behavior) %>% summarise(Average=mean(Observations), SE=se(Observations))

SumStats_taxa<-dat01 %>% pivot_longer(colnames(dat01[6:ncol(dat01)]), names_to="Insect", values_to="Observations") %>% group_by(Flower, Time, Insect) %>% summarise(Average=mean(Observations), SE=se(Observations))

taxa<-colnames(dat01[6:ncol(dat01)])
beh_lower_dat01yime<-SumStats_behavior$Average-SumStats_behavior$SE
beh_upper_daytime<-SumStats_behavior$Average+SumStats_behavior$SE



Daytime_beh<-ggplot(data=dat01 %>% pivot_longer(colnames(dat01[6:ncol(dat01)]), names_to="Insect", values_to="Observations") %>% group_by(Flower, Behavior) %>% filter(Time=="Day") %>% summarise(Average=mean(Observations), SE=se(Observations)), aes(x=Flower, y=Average, fill=Behavior))+
  geom_col(position="dodge")+
  labs(title="Average Number of Observed Insect Behaviors by Flower: Daytime", x="Flower Type", y="Average Number of Observations") 

Daytime_beh


Evening_beh<-ggplot(data=dat01 %>% pivot_longer(colnames(dat01[6:ncol(dat01)]), names_to="Insect", values_to="Observations") %>% group_by(Flower, Behavior) %>% filter(Time=="Evening") %>% summarise(Average=mean(Observations), SE=se(Observations)), aes(x=Flower, y=Average, fill=Behavior))+
  geom_col(position="dodge")+
  labs(title="Average Number of Observed Insect Behaviors by Flower: Evening", x="Flower Type", y="Average Number of Observations") 

Evening_beh


Daytime_insects<-ggplot(data=dat01 %>% pivot_longer(colnames(dat01[6:ncol(dat01)]), names_to="Insect", values_to="Observations") %>% group_by(Flower, Insect) %>% filter(Time=="Day") %>% summarise(Average=mean(Observations), SE=se(Observations)), aes(x=Flower, y=Average, fill=Insect))+
  geom_col(position="dodge")+
  labs(title="Average Number of Observed Insects by Flower: Daytime", x="Flower Type", y="Average Number of Observations") 

Daytime_insects


Evening_insects<-ggplot(data=dat01 %>% pivot_longer(colnames(dat01[6:ncol(dat01)]), names_to="Insect", values_to="Observations") %>% group_by(Flower, Insect) %>% filter(Time=="Evening") %>% summarise(Average=mean(Observations), SE=se(Observations)), aes(x=Flower, y=Average, fill=Insect))+
  geom_col(position="dodge")+
  labs(title="Average Number of Observed Insects by Flower: Evening", x="Flower Type", y="Average Number of Observations") 

Evening_insects
```



